Analysis plan

### 2023-03-29 02. Design/MSG-P2312-Analysis Plan-1-clusters.rmd
Update MSG-P2312-Analysis Plan-1-clusters.rmd
Checked for consistent use of terms and updated based on discussion with Jake on clustering.

#Proofread for typos

@@

# CONTINUE REVIEWING TERMS FROM HERE.

# Update based on notes from conversation with Jake 3-23-23

<!-- scaling variables -->
<!--  - use rank -->
<!--  - demean and divide by std deviation -->
<!-- compare clustering results -->
<!-- check variable distribution -->
<!-- check pairwise plots of variables -->

@@

For both rules-based and cluster classification of individual staff, the
daily check-in data will be transformed so that each row is one staff
person. From the raw data, the following variables will be defined:

@@

For staff-level analysis, the team will be defined by the employee's
supervisor. Changes in supervisors will come from staff moving teams or
an existing team changing supervisors. The size of the team will be
calculated with the mode of staff with the supervisor over the
observations window. For employees with multiple supervisors, the size
of the team will be the average of the modes of team sizes for each
supervisor.

@@

Teams will be defined based on the supervisors in March 2022. The unique
identifier for teams will be the `supervisor_e_mail_address`. All
employees under those supervisors will be defined as part of the team.
Employees may join and leave a team over the observation window.

For both rules-based and cluster classification of teams, the daily
check-in data will be transformed so that each row is one team as
defined by the supervisor. Supervisors may be in multiple teams as
supervisor and subordinate. From the raw data, the following variables
will be defined:

@@

-   Proportion of staff working days onsite

-   Proportion of staff working days remote

-   Proportion of staff working days remote by the day of the week

-   Proportion of staff exiting team

-   Number of staff working days

@@

1.  **All remote** - A policy definition for staff that are 100% remote
2.  **Teleworker** - A policy definition for staff that require two days
    per pay period onsite.
3.  **Onsite Flexible** - Defined by `position_category_description` in
    the roster data
4.  **Offsite** - Defined by `position_category_description` in the
    roster data
5.  **Onsite Required** - Defined by `position_category_description` in
    the roster data
6.  **Primarily remote** - Staff who work remotely above the 95th
    percentile <!-- could be 75th -->
7.  **Mostly remote** - Staff who work remotely between the median and
    the 95th percentile <!-- could be 75th -->
8.  **Mostly onsite** - Staff who work remotely between 5th percentile
    and the median <!-- could be 25th -->
9.  **Primarily onsite** - Staff who work remotely below the 5th
    percentile <!-- could be 75th -->

@@

1.  **All remote** - All team members are100% remote
2.  **Teleworker** - All team members work two days per pay period
    onsite.
3.  **Onsite Flexible** - All team members are *Onsite Flexible* as
    defined by `position_category_description` in the roster data
4.  **Offsite** - All team members are *Offsite* as defined by
    `position_category_description` in the roster data
5.  **Onsite Required** - All team members are *Onsite Required* as
    defined by `position_category_description` in the roster data
6.  Mixed position categories - Team members are different position
    categories as as defined by `position_category_description` in the
    roster data
7.  **Primarily remote** - Teams that work remotely above the 95th
    percentile <!-- could be 75th -->
8.  **Mostly remote** - Teams that work remotely between the median and
    the 95th percentile <!-- could be 75th -->
9.  **Mostly onsite** - Teams that work remotely between 5th percentile
    and the median <!-- could be 25th -->
10. **Primarily onsite** - Teams that work remotely below the 5th
    percentile <!-- could be 75th -->

@@

All of the these variables will be scaled to reduce the effects of
outliers should they be present.

@@

-   Staff work pattern statistics:

    -   Proportion in staff time telework

    -   Proportion in staff time telework by the day of the week

    -   Proportion in staff time onsite

    -   Proportion in staff time onsite by the day of the week

    -   Correlations of by staff time between the days of the week

@@

    -   Proportion in staff time telework

    -   Proportion in staff time onsite

@@

    -   Number of staff working days

@@

-   Staff descriptive statistics

    -   Number and proportion of staff by grade

    -   Number and proportion of staff by occupational series

    -   Number and proportion of staff by position category

    -   Number and proportion of staff by GSA region

    -   Number and proportion of staff by SSO

    -   Number and proportion of staff by state

    -   Number and proportion of staff by city

    -   Number and proportion of staff by supervisor level

@@

    -   Number and proportion of staff by most frequent grade

    -   Number and proportion of staff by most frequent occupational
        series

    -   Number and proportion of staff by most frequent position
        category

    -   Number and proportion of staff by most frequent GSA region

    -   Number and proportion of staff by most frequent SSO

    -   Number and proportion of staff by most frequent state

    -   Number and proportion of staff by most frequent city

    -   Number and proportion of staff by most frequent supervisor level

The staff statistics will be calculated for the following groups:

1.  Whole cohort of staff
2.  By rule-based categories of work pattern
3.  By cluster defined categories of work pattern

@@

To understand the relationship between staff work patterns and their
supervisor, we will calculate the proportion of the staff that fall into
the same rule-based and clusters as their supervisors overall, as well
as by the above categories.

>>>>

<!-- #Proofread for typos -->

@@

@@

For both rules-based and cluster classification of employee, the daily
check-in data will be transformed so that each row is one employee. From
the raw data, the following variables will be defined:

@@

For employee analysis, the team will be defined by the employee's
supervisor. Changes in supervisors will come from employees moving teams
or an existing team changing supervisors. The size of the team will be
calculated with the mode of team members with the supervisor over the
observations window. For employees with multiple supervisors, the size
of the team will be the average of the modes of team sizes for each
supervisor.

@@

Teams will be defined based on the supervisors in March 2022. The unique
identifier for teams will be the `supervisor_e_mail_address`. All team
members under those supervisors will be defined as part of the team.
Team members may join and leave a team over the observation window.

For both rules-based and cluster classification of teams, the daily
check-in data will be transformed so that each row is one team as
defined by the supervisor. Supervisors may be in multiple teams as
supervisor and team member. From the raw data, the following variables
will be defined:

@@

-   Proportion of team members working days onsite

-   Proportion of team members working days remote

-   Proportion of team members working days remote by the day of the
    week

-   Proportion of team members exiting team

-   Number of team members working days

@@

1.  **All remote** - A policy definition for staff that are 100% remote
2.  **Teleworker** - A policy definition for staff that require two days
    per pay period onsite.
3.  **Onsite Flexible** - Defined by `position_category_description` in
    the roster data
4.  **Offsite** - Defined by `position_category_description` in the
    roster data
5.  **Onsite Required** - Defined by `position_category_description` in
    the roster data
6.  **Primarily remote** - Employees who work remotely above the 95th
    percentile <!-- could be 75th -->
7.  **Mostly remote** - Employees who work remotely between the median
    and the 95th percentile <!-- could be 75th -->
8.  **Mostly onsite** - Employees who work remotely between 5th
    percentile and the median <!-- could be 25th -->
9.  **Primarily onsite** - Employees who work remotely below the 5th
    percentile <!-- could be 75th -->

@@

1.  **All remote** - All team members are100% remote
2.  **Teleworker** - All team members work two days per pay period
    onsite.
3.  **Onsite Flexible** - All team members are *Onsite Flexible* as
    defined by `position_category_description` in the roster data
4.  **Offsite** - All team members are *Offsite* as defined byteam
    members `position_category_description` in the roster data
5.  **Onsite Required** - All team members are *Onsite Required* as
    defined by `position_category_description` in the roster data
6.  Mixed position categories - Team members are different position
    categories as as defined by `position_category_description` in the
    roster data
7.  **Primarily remote** - Teams that work remotely above the 95th
    percentile <!-- could be 75th -->
8.  **Mostly remote** - Teams that work remotely between the median and
    the 95th percentile <!-- could be 75th -->
9.  **Mostly onsite** - Teams that work remotely between 5th percentile
    and the median <!-- could be 25th -->
10. **Primarily onsite** - Teams that work remotely below the 5th
    percentile <!-- could be 75th -->

@@

<!-- # Update based on notes from conversation with Jake 3-23-23 -->

<!-- scaling variables -->

<!--  - use rank -->

<!--  - demean and divide by std deviation -->

<!-- compare clustering results -->

<!-- check variable distribution -->

<!-- check pairwise plots of variables -->

All of the these variables will be scaled to reduce the effects of
outliers should they be present. Two scaling methods will be used:

1.  Normalize - Data will have the mean that that element subtracted and
    then divided by the standard deviation, using the `scale()`
    function[^1]

2.  Rank - The rank of each element will be used instead of the element.

[^1]: [Documentation for the scale
    function.](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/scale)

The scaled elements will be compared to the true data by reviewing their
distributions and comparing the results of the clusters.

@@

-   Employee work pattern statistics:

    -   Proportion in employee time telework

    -   Proportion in employee time telework by the day of the week

    -   Proportion in employee time onsite

    -   Proportion in employee time onsite by the day of the week

    -   Correlations of by employee time between the days of the week

@@

    -   Proportion in team members time telework

    -   Proportion in team members time onsite

@@

    -   Number of team members working days

@@

-   Employee descriptive statistics

    -   Number and proportion of employees by grade

    -   Number and proportion of employees by occupational series

    -   Number and proportion of employees by position category

    -   Number and proportion of employees by GSA region

    -   Number and proportion of employees by SSO

    -   Number and proportion of employees by state

    -   Number and proportion of employees by city

    -   Number and proportion of employees by supervisor level

@@

    -   Number and proportion of team members by most frequent grade

    -   Number and proportion of team members by most frequent
        occupational series

    -   Number and proportion of team members by most frequent position
        category

    -   Number and proportion of team members by most frequent GSA
        region

    -   Number and proportion of team members by most frequent SSO

    -   Number and proportion of team members by most frequent state

    -   Number and proportion of team members by most frequent city

    -   Number and proportion of team members by most frequent
        supervisor level

The employees statistics will be calculated for the following groups:

1.  Whole cohort of employees
2.  By rule-based categories of work pattern
3.  By cluster defined categories of work pattern

@@

To understand the relationship between employees work patterns and their
supervisor, we will calculate the proportion of the employees that fall
into the same rule-based and clusters as their supervisors overall, as
well as by the above categories.
